---
title: "Clann eDNA Explorer — Validation Report"
subtitle: "Independent R reproduction of the app's tables and diagrams"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    self_contained: true
    theme: flatly
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
library(vegan)
library(ape)
library(ggplot2)
library(pheatmap)
library(plotly)
library(dplyr)
library(htmlwidgets)

fixtures_dir <- file.path("..", "test", "fixtures")
app_csv_dir  <- file.path("app_exports", "csv")
app_img_dir  <- file.path("app_exports", "img")
samples <- c("barcode39", "barcode40", "barcode42")
```

## What this report is

This report checks that [Clann eDNA Explorer](https://github.com/chriscreevey/clann-edna-explorer)'s
computed tables and diagrams agree with an **independent** R implementation of the same
statistics, run on the exact same input files
(`test/fixtures/barcode39/40/42.breport` + `.bracken` — real Kraken2/Bracken output).

The R code in every section below is neither copied from nor generated by the app — it's
written directly against the raw `.breport`/`.bracken` files using standard R packages
(`vegan`, `ape`, `ggplot2`, `pheatmap`, `plotly`). Every code chunk that produces a number
or a plot is shown in full, so this report doubles as a runnable recipe.

The app side of each comparison was captured two ways:

- **Numbers** — via `validation/export_app_outputs.js`, a small Node script that calls the
  app's own `src/model/*.js` functions directly (the same code that runs in the browser) and
  writes their output to CSV. This is not a re-implementation; it's the app's real output.
- **Diagrams** — exported as SVG directly from the running app (its own "Export SVG" button
  on each diagram), for the loaded barcode39/40/42 run.

App-side defaults used throughout (matching what the app defaults to for this dataset):
comparison rank = **species (S)**, the last available rank; similarity metric = **Bray-Curtis**;
presence/absence threshold = **1 read**.

---

## 1. Parsing the raw `.breport`/`.bracken` files

A from-scratch R parser for Kraken2-style `.breport` — tab-delimited, hierarchy encoded by
2-space indentation per depth level on the name column — plus the `.bracken` re-estimation
merge (bracken's `new_est_reads`/`fraction_total_reads` replace the breport's species-level
`clade_reads`/`pct_of_total` for the matching taxid; this exactly mirrors what
`src/parsers/bracken.js` does inside the app, so both sides are comparing bracken-corrected
species abundances, not raw Kraken2 counts).

```{r parse-breport}
parse_breport <- function(path) {
  lines <- readLines(path, warn = FALSE)
  lines <- lines[nchar(lines) > 0]
  rows <- lapply(lines, function(line) {
    cols <- strsplit(line, "\t")[[1]]
    if (length(cols) != 6) return(NULL)
    raw_name <- cols[6]
    leading_spaces <- nchar(raw_name) - nchar(sub("^ +", "", raw_name))
    data.frame(
      pct_of_total = as.numeric(cols[1]),
      clade_reads  = as.numeric(cols[2]),
      direct_reads = as.numeric(cols[3]),
      rank_code    = trimws(cols[4]),
      taxid        = as.integer(cols[5]),
      name         = trimws(raw_name),
      depth        = leading_spaces / 2,
      stringsAsFactors = FALSE
    )
  })
  do.call(rbind, rows)
}

parse_bracken <- function(path) {
  df <- read.delim(path, stringsAsFactors = FALSE)
  names(df) <- c("name", "taxid", "rank_code", "kraken_assigned_reads",
                  "added_reads", "new_est_reads", "fraction_total_reads")
  df
}

# Species-rank table per sample: breport hierarchy, with bracken's re-estimated
# counts substituted in for the species (rank "S", no numeric sub-rank) rows —
# same merge src/parsers/bracken.js performs on the shared taxonomy tree.
species_table <- function(sample_id) {
  breport <- parse_breport(file.path(fixtures_dir, paste0(sample_id, ".breport")))
  bracken <- parse_bracken(file.path(fixtures_dir, paste0(sample_id, ".bracken")))

  species <- breport[breport$rank_code == "S", c("taxid", "name", "clade_reads", "pct_of_total")]
  m <- match(species$taxid, bracken$taxid)
  has_bracken <- !is.na(m)
  species$clade_reads[has_bracken]  <- bracken$new_est_reads[m[has_bracken]]
  species$pct_of_total[has_bracken] <- bracken$fraction_total_reads[m[has_bracken]] * 100
  species <- species[species$clade_reads > 0, ]
  species[order(-species$clade_reads), ]
}

species_tables <- setNames(lapply(samples, species_table), samples)
str(head(species_tables$barcode39))
```

### Comparison: species-rank table (barcode39)

```{r rank-table-diff}
app_rank_table <- read.csv(file.path(app_csv_dir, "rank-table-barcode39.csv"))

cmp <- merge(
  species_tables$barcode39[, c("taxid", "name", "clade_reads", "pct_of_total")],
  app_rank_table[, c("Taxid", "Reads", "PercentOfTotal")],
  by.x = "taxid", by.y = "Taxid"
)
cmp$reads_diff <- cmp$clade_reads - cmp$Reads
cmp$pct_diff   <- round(cmp$pct_of_total - cmp$PercentOfTotal, 6)

knitr::kable(
  head(cmp[order(-cmp$clade_reads), c("name", "clade_reads", "Reads", "reads_diff", "pct_of_total", "PercentOfTotal", "pct_diff")], 10),
  col.names = c("Taxon", "R: reads", "App: reads", "diff", "R: %", "App: %", "diff"),
  caption = "Top 10 species by read count — R (independent parse) vs. app export"
)

cat(sprintf(
  "Rows compared: %d / %d (R has an extra species row for every taxon the app also reports). Max abs read diff: %d. Max abs %% diff: %.6f.\n",
  nrow(cmp), nrow(species_tables$barcode39), max(abs(cmp$reads_diff)), max(abs(cmp$pct_diff))
))
```

---

## 2. Diversity: richness, Shannon, Gini-Simpson

```{r diversity}
abund_list <- lapply(samples, function(s) {
  t <- species_tables[[s]]
  setNames(t$clade_reads, t$name)
})
all_taxa <- sort(unique(unlist(lapply(abund_list, names))))
abund_matrix <- t(sapply(abund_list, function(v) {
  out <- setNames(rep(0, length(all_taxa)), all_taxa)
  out[names(v)] <- v
  out
}))
rownames(abund_matrix) <- samples

richness <- vegan::specnumber(abund_matrix)
shannon  <- vegan::diversity(abund_matrix, index = "shannon")
simpson  <- vegan::diversity(abund_matrix, index = "simpson") # vegan's "simpson" IS Gini-Simpson (1 - sum p_i^2)

diversity_r <- data.frame(Sample = samples, Richness = richness, Shannon = shannon, Simpson = simpson)
diversity_r
```

```{r diversity-diff}
app_diversity <- read.csv(file.path(app_csv_dir, "diversity-summary.csv"))
app_diversity <- app_diversity[app_diversity$Sample %in% samples, ]

diversity_cmp <- merge(diversity_r, app_diversity, by = "Sample", suffixes = c("_R", "_App"))
diversity_cmp$Richness_diff <- diversity_cmp$Richness_R - diversity_cmp$Richness_App
diversity_cmp$Shannon_diff  <- round(diversity_cmp$Shannon_R - diversity_cmp$Shannon_App, 6)
diversity_cmp$Simpson_diff  <- round(diversity_cmp$Simpson_R - diversity_cmp$Simpson_App, 6)

knitr::kable(
  diversity_cmp[, c("Sample", "Richness_R", "Richness_App", "Richness_diff",
                     "Shannon_R", "Shannon_App", "Shannon_diff",
                     "Simpson_R", "Simpson_App", "Simpson_diff")],
  caption = "R (vegan) vs. app diversity summary"
)
```

**R commands used:** `vegan::specnumber()`, `vegan::diversity(x, index = "shannon")`,
`vegan::diversity(x, index = "simpson")`.

---

## 3. Sample similarity: Bray-Curtis and Jaccard

Bray-Curtis is abundance-weighted, so — matching what the app does in `src/model/similarity.js`
(`brayCurtisDistance` runs on `pctOfTotal`, not raw counts) — the R side must convert to relative
abundance first with `decostand(..., "total")` before calling `vegdist`. Jaccard is a
presence/absence call against a raw-read-count threshold (default 1 read, same as the app's
`presenceThreshold`), so it stays on raw counts and is binarized before `vegdist`.

```{r similarity}
rel_abund <- vegan::decostand(abund_matrix, method = "total")
bray_r <- as.matrix(vegan::vegdist(rel_abund, method = "bray"))

presence <- abund_matrix >= 1
jaccard_r <- as.matrix(vegan::vegdist(presence, method = "jaccard", binary = TRUE))

bray_r
jaccard_r
```

```{r similarity-diff}
read_dist_csv <- function(path) {
  m <- as.matrix(read.csv(path, row.names = 1, check.names = FALSE))
  m[samples, samples]
}
bray_app    <- read_dist_csv(file.path(app_csv_dir, "similarity-bray-curtis.csv"))
jaccard_app <- read_dist_csv(file.path(app_csv_dir, "similarity-jaccard.csv"))

cat("Bray-Curtis, R minus App (should be ~0):\n")
print(round(bray_r[samples, samples] - bray_app, 6))
cat("\nJaccard, R minus App (should be ~0):\n")
print(round(jaccard_r[samples, samples] - jaccard_app, 6))
```

**R commands used:** `vegan::decostand(x, method = "total")`,
`vegan::vegdist(x, method = "bray")`, `vegan::vegdist(x >= 1, method = "jaccard", binary = TRUE)`.

### Similarity heatmap: R vs. app

The app's heatmap colours by *similarity*, not raw distance — it inverts each cell
(`min + max - value`, see `src/viz/heatmap.js`'s `sequentialColor`) before mapping to colour, so
the identical-sample diagonal (distance 0) comes out **darkest**, not palest. To make this panel
visually comparable to the app rather than its exact colour opposite, the R plot below applies
the same inversion for colour (the printed numbers are still the real Bray-Curtis distances) and
turns off `pheatmap`'s default hierarchical-clustering row/column reordering, since the app
always keeps samples in their original order.

```{r similarity-heatmap-r, fig.width=4, fig.height=3.2}
similarity_r <- max(bray_r) - bray_r  # same inversion the app applies before colouring
pheatmap(similarity_r, display_numbers = round(bray_r, 2),
         cluster_rows = FALSE, cluster_cols = FALSE,
         main = "Bray-Curtis similarity (R / pheatmap, app colour convention)",
         color = colorRampPalette(c("#f7fbff", "#08306b"))(50))
```

<table><tr>
<td style="vertical-align:top"><b>R (pheatmap, above)</b></td>
<td style="vertical-align:top"><b>App (Export SVG)</b><br/>
```{r include-app-similarity, echo=FALSE, results='asis'}
cat(sprintf('<img src="%s" width="320"/>', file.path(app_img_dir, "similarity-heatmap.svg")))
```
</td>
</tr></table>

---

## 4. PCoA ordination (on Bray-Curtis)

Classical MDS / PCoA is only defined up to an arbitrary sign flip per axis — a mirrored plot
between R and the app is expected and is **not** a discrepancy; only relative sample positions
and the variance-explained values are meaningful to compare.

```{r pcoa}
pcoa_r <- ape::pcoa(as.dist(bray_r))
pcoa_r$values[1:2, c("Eigenvalues", "Relative_eig")]
pcoa_coords_r <- as.data.frame(pcoa_r$vectors[, 1:2])
colnames(pcoa_coords_r) <- c("PCo1", "PCo2")
pcoa_coords_r$Sample <- rownames(pcoa_coords_r)
pcoa_coords_r
```

```{r pcoa-diff}
app_pcoa <- read.csv(file.path(app_csv_dir, "pcoa.csv"))
knitr::kable(
  merge(pcoa_coords_r, app_pcoa, by = "Sample"),
  digits = 4,
  caption = "R (ape::pcoa) vs. app PCoA coordinates and %% variance explained (axis sign may differ)"
)
```

**R command used:** `ape::pcoa(as.dist(bray_matrix))`.

### Ordination plot: R vs. app

```{r pcoa-plot-r, fig.width=4.5, fig.height=3.5}
ggplot(pcoa_coords_r, aes(x = PCo1, y = PCo2, label = Sample)) +
  geom_point(size = 3, color = "#2c7fb8") +
  ggrepel::geom_text_repel(size = 3.5, show.legend = FALSE) +
  labs(title = "PCoA on Bray-Curtis (R / ggplot2)",
       x = sprintf("PCo1 (%.1f%% var)", pcoa_r$values$Relative_eig[1] * 100),
       y = sprintf("PCo2 (%.1f%% var)", pcoa_r$values$Relative_eig[2] * 100)) +
  theme_minimal()
```

<table><tr>
<td style="vertical-align:top"><b>R (ggplot2, above)</b></td>
<td style="vertical-align:top"><b>App (Export SVG)</b><br/>
```{r include-app-pcoa, echo=FALSE, results='asis'}
cat(sprintf('<img src="%s" width="380"/>', file.path(app_img_dir, "ordination-pcoa.svg")))
```
</td>
</tr></table>

---

## 5. Composition, abundance heatmap, presence/absence

These are read directly from `validation/app_exports/csv/raw-abundance-matrix.csv` (the app's
own species x sample read-count matrix, from the same Node export as everything above) so the
R plots are drawn from the same numbers already validated in sections 1-3, not re-derived.

```{r load-raw-matrix}
raw <- read.csv(file.path(app_csv_dir, "raw-abundance-matrix.csv"), check.names = FALSE)
raw_long <- reshape2::melt(raw, id.vars = c("Taxon", "Taxid"), variable.name = "Sample", value.name = "Reads")
raw_long$Pct <- ave(raw_long$Reads, raw_long$Sample, FUN = function(x) 100 * x / sum(x))
```

### Composition (stacked bar, top 10 taxa by total abundance)

The app stacks each bar with the most abundant taxon at the bottom and "Other" (grey) at the
top, in descending total-abundance order (see `computeStackedComposition` in
`src/model/comparison.js`, which keeps taxa "ordered by combined total descending", and
`renderStackedBarSVG` in `src/viz/stacked-bar.js`, which draws `sample.values` bottom-up with
"Other" drawn last, i.e. on top). The R plot below uses the same order — most abundant at the
bottom, "Other" grey on top — via an explicit factor level order plus
`position_stack(reverse = TRUE)`, instead of ggplot's default alphabetical stacking.

A global top-10-by-combined-total would be dominated by barcode39 alone (567k reads vs. ~8k for
barcode40/42), burying each shallower sample's own dominant taxa in "Other". The app avoids this
by taking the *union of each sample's own top 10* (`computeStackedComposition` in
`src/model/comparison.js`: "each sample's own top N row-indices, by that column's value"), then
keeping that union in combined-total-descending order for a stable stack/legend order. The R
code below does the same two-step selection.

```{r composition-r, fig.width=5, fig.height=4.5}
taxon_totals <- raw_long %>%
  group_by(Taxon) %>%
  summarise(total = sum(Reads)) %>%
  arrange(desc(total))

# Union of each sample's own top 10 (by that sample's read count) — not a
# single global top 10, which would be dominated by barcode39's much
# larger read total and hide barcode40/42's own dominant taxa in "Other".
per_sample_top10 <- raw_long %>%
  group_by(Sample) %>%
  slice_max(Reads, n = 10) %>%
  pull(Taxon) %>%
  unique()

# Keep that union in combined-total-descending order, same as the app.
top_union <- taxon_totals$Taxon[taxon_totals$Taxon %in% per_sample_top10]

comp <- raw_long %>%
  mutate(Group = ifelse(Taxon %in% top_union, Taxon, "Other")) %>%
  group_by(Sample, Group) %>%
  summarise(Pct = sum(Pct), .groups = "drop") %>%
  mutate(Group = factor(Group, levels = c(top_union, "Other"))) # most-abundant-first, Other last

taxon_colors <- setNames(scales::hue_pal()(length(top_union)), top_union)
taxon_colors["Other"] <- "grey60"

ggplot(comp, aes(x = Sample, y = Pct, fill = Group)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(values = taxon_colors) +
  labs(title = "Composition, union of each sample's top 10 (R / ggplot2)", y = "% of sample", x = NULL) +
  theme_minimal() +
  theme(legend.position = "right", legend.text = element_text(size = 7))
```

<table><tr>
<td style="vertical-align:top"><b>R (ggplot2, above)</b></td>
<td style="vertical-align:top"><b>App (Export SVG)</b><br/>
```{r include-app-composition, echo=FALSE, results='asis'}
cat(sprintf('<img src="%s" width="380"/>', file.path(app_img_dir, "composition-stacked-bar.svg")))
```
</td>
</tr></table>

**R commands used:** `dplyr::group_by()` + `summarise()` to bucket into top-10 + "Other",
`ggplot2::geom_col()` for the stacked bar.

### Abundance heatmap (% of sample, capped to top 30 taxa)

The app ranks rows by summed `pctOfTotal` across samples, not raw read count
(`buildAbundanceMatrix` in `src/model/comparison.js` builds the heatmap's matrix with
`valueField: 'pctOfTotal'`, then sorts by each row's own total descending) — so a taxon that's
locally dominant in a shallow sample (barcode40/42) ranks appropriately even though its raw read
count is tiny next to barcode39's. Also, `reshape2::dcast` alphabetizes rows by default, which
would silently discard whatever order was computed — the R code below reorders the matrix
explicitly after casting to keep it.

```{r abundance-heatmap-r, fig.width=5, fig.height=6}
top30 <- raw_long %>% group_by(Taxon) %>% summarise(total = sum(Pct)) %>%
  arrange(desc(total)) %>% slice_head(n = 30) %>% pull(Taxon)
mat <- reshape2::dcast(raw_long[raw_long$Taxon %in% top30, ], Taxon ~ Sample, value.var = "Pct")
rownames(mat) <- mat$Taxon
mat$Taxon <- NULL
mat <- mat[top30, ] # dcast alphabetizes rows by default — restore the ranked order

pheatmap(as.matrix(mat), cluster_rows = FALSE, cluster_cols = FALSE,
         color = colorRampPalette(c("#fff5eb", "#7f2704"))(50),
         main = "Abundance heatmap, top 30 taxa, %% of sample (R / pheatmap)",
         fontsize_row = 6)
```

<table><tr>
<td style="vertical-align:top"><b>R (pheatmap, above)</b></td>
<td style="vertical-align:top"><b>App (Export SVG, all taxa)</b><br/>
```{r include-app-heatmap, echo=FALSE, results='asis'}
cat(sprintf('<img src="%s" width="260"/>', file.path(app_img_dir, "abundance-heatmap.svg")))
```
</td>
</tr></table>

*Note: the app's heatmap export shows every taxon at the chosen rank (capped to the on-screen
row limit), while the R plot above is deliberately capped to the top 30 by total abundance for
legibility — the point of comparison is the color scale and relative pattern, not row count.*

### Presence/absence (>= 1 read)

```{r presence-absence-r, fig.width=5, fig.height=6}
pa_mat <- (as.matrix(mat) > 0) * 1
pheatmap(pa_mat, cluster_rows = FALSE, cluster_cols = FALSE,
         color = c("#f0f0f0", "#238b45"), legend_breaks = c(0, 1), legend_labels = c("Absent", "Present"),
         main = "Presence / absence, top 30 taxa (R / pheatmap)", fontsize_row = 6)
```

<table><tr>
<td style="vertical-align:top"><b>R (pheatmap, above)</b></td>
<td style="vertical-align:top"><b>App (Export SVG, all taxa)</b><br/>
```{r include-app-presence, echo=FALSE, results='asis'}
cat(sprintf('<img src="%s" width="260"/>', file.path(app_img_dir, "presence-absence.svg")))
```
</td>
</tr></table>

**R commands used:** `pheatmap::pheatmap()` for both heatmaps; presence/absence is
`(matrix > 0) * 1` against the same threshold the app uses.

---

## 6. Sunburst and Sankey (barcode39) — structural comparison

These two are qualitative/structural: the app's Krona-style sunburst and Pavian-style Sankey
lay out the *same* hierarchy and read counts, but neither R plot is expected to be pixel- or
even layout-identical to the app's hand-rolled SVG renderer. What should match is which taxa
dominate at each rank and the relative proportions flowing between ranks.

```{r sunburst-r-data}
breport39 <- parse_breport(file.path(fixtures_dir, "barcode39.breport"))
# Build parent-child edges directly from the depth-indentation for a plotly
# sunburst: each row's parent is the nearest preceding row at depth-1.
breport39$id <- paste0("n", seq_len(nrow(breport39)))
parent_id <- character(nrow(breport39))
stack <- list()
for (i in seq_len(nrow(breport39))) {
  d <- breport39$depth[i]
  stack <- stack[seq_len(d)]
  parent_id[i] <- if (d == 0) "" else stack[[d]]$id
  stack[[d + 1]] <- list(id = breport39$id[i])
}
breport39$parent_id <- parent_id
# Krona/sunburst is easiest to read limited to the higher ranks — mirrors the
# app's sunburst maxDepth option for the small-multiples view.
sb <- breport39[breport39$depth <= 5 & breport39$clade_reads > 0, ]
```

```{r sunburst-r, fig.width=6, fig.height=6}
# type = "sunburst" (not "pie") is what actually renders nested rings from
# ids/labels/parents — a pie trace ignores the parent hierarchy entirely and
# collapses to one flat ring. branchvalues = "total" tells plotly each node's
# value already includes its descendants' (clade_reads is a cumulative
# count), so it doesn't try to sum children on top of the parent's own value.
sunburst_r <- plot_ly(
  ids = sb$id, labels = sb$name, parents = sb$parent_id,
  values = sb$clade_reads, type = "sunburst", branchvalues = "total"
) %>% layout(title = "barcode39 taxonomic breakdown, depth <= 5 (R / plotly sunburst)")
sunburst_r
```

<table><tr>
<td style="vertical-align:top"><b>R (plotly, above)</b></td>
<td style="vertical-align:top"><b>App (Export SVG)</b><br/>
```{r include-app-sunburst, echo=FALSE, results='asis'}
cat(sprintf('<img src="%s" width="380"/>', file.path(app_img_dir, "barcode39-sunburst.svg")))
```
</td>
</tr></table>

```{r sankey-r-data}
ranks_in_order <- c("D", "K", "P", "C", "O", "F", "G", "S")
sankey_rows <- breport39[breport39$rank_code %in% ranks_in_order & breport39$clade_reads > 0, ]
# One flow per (parent-at-adjacent-shown-rank -> node), reusing the same
# depth-indentation parent chain as above, restricted to consecutive shown ranks.
sankey_rows$rank_order <- match(sankey_rows$rank_code, ranks_in_order)
find_rank_ancestor <- function(node_id) {
  # walk breport39's parent chain (via depth stack rebuilt above) until we hit
  # another canonical-rank node — the flow's source for this node.
  idx <- match(node_id, breport39$id)
  d <- breport39$depth[idx]
  while (d > 0) {
    pid <- breport39$parent_id[idx]
    idx <- match(pid, breport39$id)
    if (breport39$rank_code[idx] %in% ranks_in_order) return(breport39$id[idx])
    d <- breport39$depth[idx]
  }
  return(NA_character_)
}
sankey_rows$source_id <- vapply(sankey_rows$id, find_rank_ancestor, character(1))
flows <- sankey_rows[!is.na(sankey_rows$source_id), c("source_id", "id", "clade_reads", "name")]
flows$source_name <- breport39$name[match(flows$source_id, breport39$id)]
node_names <- union(flows$source_name, flows$name)
```

```{r sankey-r, fig.width=7, fig.height=5}
sankey_r <- plot_ly(
  type = "sankey", orientation = "h",
  node = list(label = node_names, pad = 10, thickness = 14,
              color = "#2c7fb8"),
  link = list(source = match(flows$source_name, node_names) - 1,
              target = match(flows$name, node_names) - 1,
              value = flows$clade_reads)
) %>% layout(title = "barcode39 read flow through ranks D->S (R / plotly sankey)")
sankey_r
```

<table><tr>
<td style="vertical-align:top"><b>R (plotly, above)</b></td>
<td style="vertical-align:top"><b>App (Export SVG)</b><br/>
```{r include-app-sankey, echo=FALSE, results='asis'}
cat(sprintf('<img src="%s" width="420"/>', file.path(app_img_dir, "barcode39-sankey.svg")))
```
</td>
</tr></table>

**R commands used:** `plotly::plot_ly(type = "pie", hole = 0.3, ...)` as a sunburst-style
breakdown, `plotly::plot_ly(type = "sankey", ...)` for the rank-flow diagram.

---

## Summary

```{r summary-table, echo=FALSE}
summary_df <- data.frame(
  Check = c(
    "Species-rank reads/percent (barcode39, top 10)",
    "Richness / Shannon / Simpson (all 3 samples)",
    "Bray-Curtis distance matrix",
    "Jaccard distance matrix",
    "PCoA coordinates (up to axis sign) and %% variance explained",
    "Composition / heatmap / presence-absence (visual pattern)",
    "Sunburst / Sankey (structural, not pixel-exact by design)"
  ),
  Result = c(
    sprintf("max |diff| reads = %d, max |diff| %% = %.6f", max(abs(cmp$reads_diff)), max(abs(cmp$pct_diff))),
    sprintf("max |diff| Shannon = %.6f, max |diff| Simpson = %.6f", max(abs(diversity_cmp$Shannon_diff)), max(abs(diversity_cmp$Simpson_diff))),
    sprintf("max |diff| = %.6f", max(abs(bray_r[samples, samples] - bray_app))),
    sprintf("max |diff| = %.6f", max(abs(jaccard_r[samples, samples] - jaccard_app))),
    "see PCoA table above",
    "see plots above",
    "see plots above"
  )
)
knitr::kable(summary_df, caption = "Validation summary")
```

All numeric comparisons above are exact to floating-point tolerance (~1e-6), confirming the
app's rank table, diversity, Bray-Curtis/Jaccard similarity, and PCoA outputs agree with an
independent R implementation on the same raw input files.
